MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
Added Figure 9, updated figure 8
local update histogram pdf

\section {User Evaluation}
\label {sec:user_eval}
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth,  trim = 0mm 00mm 0mm 0mm, clip]{images_pvis/usereval2}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap. The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context". The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree. }
	\label{fig:userstudy}
\end{figure}
Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work. We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques. LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree". Fig.~\ref{fig:comparison} shows the results. Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A). This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user. As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone. Another common approach is to color map the top eigenvector of Hessian matrix to RGB space. We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query. Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before. The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective. Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E). All participants could correctly identify the number of fiber bundles. All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering. They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts. We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset. Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters. The unit for length is the grid cube length. Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Sec.~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand, a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests. Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited. Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}[tb]
%	\centering
%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
%	\label{fig:comparison}
%\end{figure}
%

>>>>

\section {User Evaluation}
\label {sec:user_eval}
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth,  trim = 0mm 30mm 0mm 0mm, clip]{images_pvis/usereval.pdf}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{The user evaluation consists of a set of queries on videos of our results, volume rendering and Hessian colormap. The questionnaire tests the effectiveness of MetaTracts in visualizing ``geometric structure" and ``spatial context". The Likert scale goes from 1-5: strongly disagree, disagree, neutral, agree and strongly agree. }
	\label{fig:userstudy}
\end{figure}
Four NDT practitioners who are familiar with X-Ray computed tomography and its visual representations, four material scientists who are experienced in analyzing fiber reinforced polymers and four practitioners of visualization participated in the user evaluation of our work. We therefore setup a questionnaire with questions on solving given tasks as well as questions using Likert scale (LS) for evaluating the performance of the MetaTracts over currently used techniques. LS consists of 5 levels from ``Strongly disagree" to ``Strongly agree". Fig.~\ref{fig:comparison} shows the results. Task 1 of the questionnaire was to ``identify the number of fiber bundles" from a video showing a 360 degree rotation of the volume rendering of dataset 1 (Fig.~\ref{fig:data-char}A). This task tests the perception of geometric structure and how common problems like uniform gray scale values and occlusion of the different fiber bundles affects the user. As expected, Fig.~\ref{fig:comparison} shows that the participants found it difficult to determine the number of bundles from volume rendering alone. Another common approach is to color map the top eigenvector of Hessian matrix to RGB space. We show a video showing a 360 degree rotation of the volume rendering of the color mapped eigenvectors (similar to Fig.~\ref{fig:reliable_hessian}A) and asked the same query. Compared to the volume rendering, the Hessian color map provides more information about the major orientations and we see that the evaluators performed better in this test than in the direct volume rendering test before. The majority of our participants agreed or strongly agreed (LS) that Hessian color map was more effective. Finally, we showed the MetaTracts extracted bundles (for orientation 1 and 2 separately Fig.~\ref{fig:orientation_clustering}D,E). All participants could correctly identify the number of fiber bundles. All participants agreed that MetaTracts was more effective in perceiving geometric structure than direct volume rendering. They especially liked the 2D slices of the MetaTracts (Fig.~\ref{fig:crop-16-decomp}B, compared to Fig.~\ref{fig:data-char}C).
The next set of tasks on the MetaTracts results tests the spatial context perceived from MetaTracts. We achieve this by asking the evaluators to select the largest fiber bundle from a subset of fiber bundles, to select the smallest from a subset and given a fiber bundle and to select the most similar one from a subset. Due to the problems of occlusions and similar gray scale values these tests are difficult on volume rendering of the plain data only, but straight forward using MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority participants agreed that the extracted meshes provides better spatial context than volume rendering.

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth,  trim = 0mm 0mm 10mm 00mm, clip]{images_pvis/histogram.pdf}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters. The unit for length is the grid cube length. Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Sec.~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand, a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests. Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited. Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}[tb]
%	\centering
%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
%	\label{fig:comparison}
%\end{figure}
%
